---
title: "ADVR_project"
author: "Varun Kumar"
date: "7/29/2020"
output: html_document
runtime: shiny
---

```{r, libraries,message=FALSE,warning=FALSE}
#Including the necessay libraries
rm(list = ls())
library(readr)
library(dplyr)
library(magrittr)
library(ggplot2)
library(shiny)
library(rstan)
library(shinystan)
library(ggiraph)
library(gganimate)
library(reshape2)
library(tidyr)
library(ggmap)
library(lubridate)
```
First we Import the exo planet data from the CSV file after setting the working directory.
```{r, warning=FALSE, message=FALSE}
#Setting the working directory
setwd("C:\\Users\\vk\\Desktop")

#Importing the data
exo <- read_csv("exo_data.csv")
#viewing the structure of the data
str(exo)
```



converting the data as per the question, We see that the columns 1,16,17,18,25 are already in character format and rest need to be converted. We do this using dplyr and magrittr package
```{r}
#Converting to factor
exo$flag %<>% as.factor
exo$meth %<>% as.factor

#converting to integer
exo$year %<>% as.integer

#Verifying the change
str(exo)
```

Q2) Exclude the exoplanets with an unknown method of discovery.

```{r}
#Removing the exo planets with unknown method of discovery
exo <-exo[which(!exo$meth == ""),]
```

Q3) Create a graphic which illustrates the relationship between the log-distances from the Sun and the methods of discovery.
```{r,message=FALSE,warning=FALSE}
#ggplot to show relationship between log distances and method
ggplot(exo, aes(x=log(dist), fill=meth, color=meth)) +
  geom_histogram(position="identity") +
  labs(title="Log-Distance from Sun by Method of Discovery",x="Log-Dist from Sun (parsec)", y = "Count")
```

 
Q4) Create scatter plots of the log-mass versus log-distances, separating by methods of discovery. Hovering with the cursor highlights the point and displays its name, and,if you click, the exo planet’s page on the Open Exo planet Catalogs will be opened.(paste the id after http://www.openexoplanetcatalogue.com/planet/).
```{r, message=FALSE,warning=FALSE}
exo$onclick <- sprintf("window.open(\"%s%s\")",
                        "http://www.openexoplanetcatalogue.com/planet/",
                        exo$id)

gg_graph = ggplot(exo,
                  aes(x = log(mass),
                      y = log(dist),
                      color = meth)) +
                  xlab('Log(Mass)') +
                  ylab('Log(Dist') +
                  scale_color_discrete(name="Discovery Method")+
                  geom_point_interactive(aes(exo_id = id,
                             tooltip = id,
                             onclick = onclick)) +
                  labs(title="Scatterplots of log-mass vs. log-distances")

ggiraph(code = print(gg_graph))
```

Q5) Rename the radius into jupiter_radius, and create a new column called earth_radius which is 11.2 times the Jupiter radius.
```{r}
#renaming using the tidyverse package function
exo <- exo %>% 
          rename(jupiter_radius = radius ) 
#creating new column as per the question
exo <- exo %>%
          mutate(earth_radius = jupiter_radius * 11.2 ) 
```



Q6) Focus only on the rows where log-earth radius and log-period have no missing values,and perform k means with four clusters on these two columns.

```{r}
# create new df for clustering
exo_clustering <- exo 

# taking the radius values excluding the NAs and focusing only on log-radius of Earth and log-period
exo_clustering <- exo %>% drop_na(earth_radius, period)  # 2732 obs

exo_clustering <- exo_clustering %>%
                      mutate(LogERadius = log(earth_radius),
                             LogPeriod  = log(period))


exo_kmeans <- exo_clustering %>%
                      select(LogERadius,LogPeriod)

# performing k-means with 4 clusters
clust_km4  <- kmeans(exo_kmeans, 4)

#displaying the cluster result
table(clust_km4$cluster)
```


Q7) Add the clustering labels to the data set through a new factor column called type,with levels rocky,hot_jupiters,cold_gas_giants,others; similarly tohttps://en.wikipedia.org/wiki/Exoplanet#/media/File:ExoplanetPopulations-20170616.pngand produce the scatter plot highlighting these clusters.
```{r}
clust_km4$cluster <- as.factor(clust_km4$cluster)

ggplot(exo_kmeans, aes(LogPeriod,LogERadius ,color = clust_km4$cluster)) + geom_point() +
                                      labs(title="Clustering solutions of Exoplanets")
```
The cluster 1,2,3,4 are cold_gas_giants, others, Rock & hot_jupiters respectively.
```{r}
exo_clustering$type <- clust_km4$cluster
exo_clustering$type <- as.numeric(exo_clustering$type)

exo_clustering$type[exo_clustering$type == 1] <- "cold_gas_giants"
exo_clustering$type[exo_clustering$type == 2] <- "others"
exo_clustering$type[exo_clustering$type == 3] <- "Rocky"
exo_clustering$type[exo_clustering$type == 4] <- "hot_jupiters"

#Checking if the headers are assigned
table(exo_clustering$type) 
```


Q8) Use a violin plot to illustrate how these clusters relate to the log-mass of the exo planet

In order to see how the clusters relate to the log-mass, we can make use of a histogram or a violin plot.
```{r, message=FALSE, warning=FALSE}
#Histogram
ggplot(exo_clustering, aes(x = log(mass))) +
                            geom_histogram(aes(color = type, fill = type), 
                                            position = "identity", bins = 30, alpha = 0.4) +
                                                       labs(title="log mass of the exoplanet W.r.t Clusters")
```


```{r,message=FALSE, warning=FALSE}
#Violin Plot
ggplot(exo_clustering, aes(x = type, y = log(mass))) + 
  geom_violin() + labs(title="Violin plot of log-mass of the exoplanet")
```



Q9) Transform r_asc and decl into two new variables that are the same variables but in values of seconds. Use these as coordinates to represent a celestial map for the exo planets.

```{r, message=FALSE, warning=FALSE}
# converting r_asc to hh:mm:ss format
exo$r_asc <- gsub(" ",":", exo$r_asc, fixed=TRUE) 
exo$r_asc <- hms(exo$r_asc) 

#storing in new variable
exo$r_asc_second <- period_to_seconds(exo$r_asc)

# convert Declination to hh:mm:ss format
exo$decl <- gsub(" ", ":", exo$decl, fixed=TRUE) 
exo$decl <- hms(exo$decl) 

#Storing in new variable
exo$decl_second <- period_to_seconds(exo$decl)

# scatter plot representing a celestial map for the exo planets
ggplot(exo, aes(r_asc_second, decl_second, color= meth)) +
                            geom_point() + 
                        labs(title="Celestial Map for Exo planets", x="Right ascension (seconds)",y="Declination (seconds)")
```


Q10) Create an animated time series where multiple lines illustrate the evolution overtime of the total number of exo planets discovered for each method up to that year
```{r, message=FALSE, warning=FALSE}
exo.timeseries <- exo %>% group_by(meth, year) %>%  summarise(Count = length(meth)) %>%
                          mutate(Count = cumsum(Count))

exo.timeseries <- na.omit(exo.timeseries)

#Animated Time series
ggplot(exo.timeseries, aes(x = year, y = Count, group = meth)) + 
  geom_line(aes(color = meth)) + 
  geom_point(size = 2) + 
  transition_reveal(year) + 
  labs(title = 'Evolution of exoplanets by methods', y = 'Number Discovered') 
```



Q11) Create an interactive plot with Shiny where you can select the year (slider widget,with values≥2009) and exo planet type. Exo planets appear as points on a scatter-plot (log-mass vs log-distance coloured by method) only if they have already been discovered. If type is equal to 'all' all types are plotted together.


```{r, echo=TRUE, message=FALSE, warning=FALSE}

# Defining UI for histogram app
ui <- fluidPage(
    
        
            sliderInput("slider",
                        "Select Year", min=2009,
                        max(exo_clustering$year,na.rm = TRUE),max(exo_clustering$year,na.rm = TRUE),
                        round = TRUE, sep = '', width = '100%',step = 1,ticks = FALSE)
        ,
        selectInput(inputId = "exo_type",
                    label = "Type",
                    choices = c('hot_jupiters','Rocky','cold_gas_giants','others',"all"),
                    selected = 0),
        
       
      
        

# Show a plot of the generated distribution
        
plotOutput("scatter1",height = 450)
        
    
)

# Defining server logic for histogram
server <- function(input, output) 
{
  #Scatter plot of all discoveries by years of planets with known mass
output$scatter1 = renderPlot({
      if(input$exo_type=="all")
      {
        scatter_1<-exo_clustering %>%filter(., year <= input$slider, !is.na(mass),
                                     !is.na(dist))
      }
      else
      {
        scatter_1<-exo_clustering %>%filter(., year <= input$slider, !is.na(mass),
                                     !is.na(dist)) %>% filter(., type == input$exo_type)
      }
      
        
        ggplot(scatter_1, aes(x = log(dist), y = log(mass),
                                 color = meth)) +
            geom_point(size = 3) +
            labs(title = paste("Discovery Year", input$slider),
                 x = "Distance of exo planet from sun (Parsecs)",
                 y = "Planet Mass (Jupiter Mass)",
                 colour = NULL) 
            
    })
  
}

#Running the app
shinyApp(ui = ui, server = server)


```


Q12) Fit a linear regression model where log period is the response variable and the logs of host_mass,host_temp and axis are the covariates (exclude rows that contain at least one missing value). Include an intercept term in the regression model.

```{r}
#Creating new data for regression analysis
analysis_exo <- exo[,c("period","host_mass","host_temp","axis")] 
#Data with missing values are omitted
analysis_exo.complete <- na.omit(analysis_exo)
#We fit the linear model 
fit<-lm(log(period)~log(host_mass)+log(host_temp)+log(axis),data = analysis_exo)
summary(fit)
```



Q13) Include in your R Markdown document some model summaries and an interpretation of the model you have fit.

From the coefficients and estimates, 
- We see that the estimate value is 5.65 and the p value is less than 0.05
- The covariate becomes significant when p value is less than 0.05.
- With unit increase in log hot star mass we see that the period decreases by 0.412, unit change in log hot star temperature we see that the period increases by 0.02 and finally unit increase in hot star mass we see that the number of days increases to 1.488.

From the Residual Summary statistics,
- We know that the median should be close to zero which is the same in our case.

Considering the hypothesis,
- Here we can reject the null hypothesis that the coefficient is 0 as the probability is very low
- The hot star mass covariate plays an important role in predicting and building the model.
- Also from the p value we can say that axis plays an important role too.
    
Considering the adjusted R squared value,
- From the adjusted R square value we can say that 99.3% of the variation can be explained by the predictor variables.



Q14) Embed the Shiny app from (11) in your RMarkdown document.
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Defining UI for histogram app
ui <- fluidPage(
    
        
            sliderInput("slider",
                        "Select Year", min=2009,
                        max(exo_clustering$year,na.rm = TRUE),max(exo_clustering$year,na.rm = TRUE),
                        round = TRUE, sep = '', width = '100%',step = 1,ticks = FALSE)
        ,
        selectInput(inputId = "exo_type",
                    label = "Type",
                    choices = c('hot_jupiters','Rocky','cold_gas_giants','others',"all"),
                    selected = 0),
        
       
      
        

# Show a plot of the generated distribution
        
plotOutput("scatter1",height = 450)
        
    
)

# Defining server logic for histogram
server <- function(input, output) 
{
  #Scatter plot of all discoveries by years of planets with known mass
output$scatter1 = renderPlot({
      if(input$exo_type=="all")
      {
        scatter_1<-exo_clustering %>%filter(., year <= input$slider, !is.na(mass),
                                     !is.na(dist))
      }
      else
      {
        scatter_1<-exo_clustering %>%filter(., year <= input$slider, !is.na(mass),
                                     !is.na(dist)) %>% filter(., type == input$exo_type)
      }
      
        
        ggplot(scatter_1, aes(x = log(dist), y = log(mass),
                                 color = meth)) +
            geom_point(size = 3) +
            labs(title = paste("Discovery Year", input$slider),
                 x = "Distance of exo planet from sun (Parsecs)",
                 y = "Planet Mass (Jupiter Mass)",
                 colour = NULL) 
            
    })
  
}

#Running the app
shinyApp(ui = ui, server = server)

```